---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
---


## Loading and preprocessing the data
```{r loading and tranforming csv file,  echo=TRUE}
require(dplyr)

file_url <- "activity.zip"
temp_file_name <- basename(file_url)
#download.file(file_url,temp_file_name)
my.csv.file <- unzip(temp_file_name) %>% read.csv(stringsAsFactors = F) %>% mutate(date = as.POSIXct(date, format="%Y-%m-%d"))
#file.remove(temp_file_name)
head(my.csv.file,15)
```

## What is mean total number of steps taken per day?

```{r Calculating total.steps.daily, echo=TRUE}
total.steps.daily <- filter(my.csv.file,!is.na(steps)) %>%
  group_by(date) %>% summarize(steps = sum(steps)) %>%
  print

```

## What is the average daily activity pattern?
```{r Histogram of total.steps.daily, echo=TRUE}
 require(ggplot2)
ggplot(data = total.steps.daily,aes(steps)) +
        geom_histogram(binwidth = 999) +
        labs(x = "steps.daily",y="Frequency")
```
## Means and Medians of daily steps 
```{r}
mean.median.total.steps.daily <- filter(my.csv.file,!is.na(steps)) %>%
  group_by(date) %>% summarize(mean.daily.steps = mean(steps),
                               median.daily.steps = median(steps)) %>%
                               
  print
```
## Time series plot of the average number of steps taken
```{r Time series of steps taken, echo=TRUE}
avg.step <- filter(my.csv.file,!is.na(steps)) %>%
  group_by(interval) %>% summarise(steps = mean(steps))
ggplot(avg.step, aes(interval,steps, col= "Red")) + geom_line() 

```
## The 5-minute interval that, on average, contains the maximum number of steps
```{r The 5-minute interval contains the maximum number of steps, echo=TRUE}
  avg.step[which.max(avg.step$steps),]
```
## Imputing missing values
```{r Imputing missing values, echo=TRUE}
require(mice)
temp_data <- select(my.csv.file,interval,steps) %>%mice(method="sample") %>% complete()
date <- my.csv.file$date
my.csv.file.updated <- mutate(temp_data,date = date)
```

## Histogram of the total number of steps taken each day after missing values are imputed
```{r, Histogram of the total number of steps after missing values imputed, echo=TRUE}
ggplot(data = my.csv.file.updated,aes(steps)) +
        geom_histogram(binwidth = 250) +
        labs(x = "steps.daily",y="Frequency")
```
## Mean and median after missing values are inserted 

```{r Mean and Median after missing values inserted, echo=TRUE}
mean.median.total.steps.daily.complete <- group_by(my.csv.file.updated,date) %>% summarize(mean.daily.steps = mean(steps),
                               median.daily.steps = median(steps)) %>%
                print               
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r extracting weekdays and weekend from date, echo=TRUE}
week.ends <- c("Saturday","Sunday")
date.factor <- ifelse(weekdays(my.csv.file$date) %in% week.ends,"Weekends","Weekdays")
data.with.date.factor <- mutate(my.csv.file,date.factor = as.factor(date.factor))
mean.steps.daily <- filter(data.with.date.factor,!is.na(steps)) %>%
group_by(date.factor,interval) %>% summarize(steps = mean(steps)) 

ggplot(mean.steps.daily, aes(interval,steps)) + geom_line() + facet_wrap(~date.factor,nrow = 2,ncol = 1) 



```

## Are there differences in activity patterns between weekdays and weekends?
### Yes, the weekdays had more activities than the weekends
